---
title: "PS07"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# SDS 192 Mini Project 2: "Follow the Money: Analysis of 2016 Presidential Election"

## Data import
```{r}
library(fec16)
library(ggplot2)
library(dplyr)
library(ggpubr)
library(tidyverse)

# Download full contributions data from web and overwrite contributions using:
# read_all_*() function.
# You will need to install the usethis package for this line of code to work
contributions <- read_all_contributions()
expenditures <- read_all_expenditures()
transactions <- read_all_transactions()
```



## Summary
These visualizations compare firstly how contributions correlate with the number of votes received by Clinton and Trump in the 2016 election, and secondly which occupations contributed the most. Trump gained more votes and contributions in swing states, which corresponds to the 2016 election result. We also found a weak correlation between votes and contribution. Occupations of each candidate’s individual contributors reveal that since Trump attracted a larger number of unemployed people’s contribution and a larger amount of presidents’ contributions, while Clinton’s contributors did not contribute as much as Trump’s, Trump was thus more advantageous and won the 2016 election.

# Votes vs Contribution of Clinton and Trump

From the top graph in the figure 1, we can see a weak correlation between the contribution and the total amount of votes in the 2016 presidential election. Namely, if a candidate gained more votes, then probably the contribution of this candidate would be high too. Since Trump won the 2016 election, we would hypothesize that Trump gained both more votes and contributions in important areas based on the former graph. Therefore, we look at the total number of votes and total amount of contribution in the swing states ^[https://en.wikipedia.org/wiki/Swing_state], which largely determines who would win in the end, for both final candidates. As we can see on the bottom two graphs, Trump gained more on both contribution and votes in swing states, which matched the result. Especially, Trump gained almost 25% more contribution in the swing states. We thus wonder why Trump gained more money, specifically what kind of people were contributing to each candidate in the 2016 election, which is discussed in the second visualization in Figure 2.
```{r}
#Clinton id: P00003392
#Trump id: P80001571

transaction_amt_C_T <- contributions %>% 
  filter(cand_id == "P00003392" | cand_id == "P80001571") %>% 
  inner_join(candidates, by = "cand_id") %>% 
  group_by(state, cand_id) %>% 
  filter(state == "AZ" | state == "CO" | state == "FL" | state == "GA" | 
           state == "IA" | state == "MI" | state == "MN" | state == "NV"  | 
           state == "NH" ) %>% 
  summarise(amt = sum(transaction_amt)) %>% 
  mutate(Candidate = ifelse(cand_id == "P00003392", "Clinton", "Trump"))
 
Votes <- results_president %>% 
  filter(cand_id ==  "P00003392" | cand_id == "P80001571") %>% 
  mutate(Candidate = ifelse(cand_id == "P00003392", "Clinton", "Trump")) %>% 
  group_by(Candidate, state) %>% 
  filter(state == "AZ" | state == "CO" | state == "FL" | state == "GA" | 
           state == "IA" | state == "MI" | state == "MN" | state == "NV"  | 
           state == "NH" ) %>% 
  summarise(Vote = sum(general_votes))

votes <- ggplot(Votes, aes(x = Candidate, y = Vote, fill = Candidate)) +
  geom_col() +
  #facet_wrap(~state) +
  scale_fill_manual(values = c("blue", "red")) + 
  labs(title = "Total Votes in swing states",
       y = "Votes")+
  theme(plot.title = element_text(size = 10))+
  scale_y_continuous(labels = scales::comma)

Vcontribution<-ggplot(transaction_amt_C_T, aes(x = Candidate, 
                                               y = amt, 
                                               fill = Candidate)) + 
  geom_col() +
  theme(axis.text.x = element_text(size = 4)) +
  scale_fill_manual(values = c("blue","red")) +
  labs(title = "Total amount of contributions in swing states ", 
       y = "Total amount (dollars)")+
  theme(plot.title = element_text(size=10))+
  scale_y_continuous(labels = scales::comma)


C_T <- candidates %>% 
  filter(cand_id == "P00003392" | cand_id == "P80001571")

results_C_T <- results_president %>%
  filter(cand_id == "P00003392" | cand_id == "P80001571") %>% 
  select(cand_id, state, general_votes, general_percent) %>% 
  arrange(cand_id)
 
contributions_C_T <- contributions %>% 
  filter(cand_id == "P00003392" | cand_id == "P80001571") %>% 
  group_by(cand_id, state) %>% 
  summarise(amt = sum(transaction_amt)) %>% 
  arrange(cand_id)

results_and_contribution_C_T <- results_C_T %>% 
  inner_join(contributions_C_T, by = c("state", "cand_id")) %>% 
  inner_join(C_T %>% select(cand_id, cand_name), by="cand_id") %>% 
  mutate(Candidate = ifelse(cand_id == "P00003392", "Clinton", "Trump")) 

VS<-ggplot(results_and_contribution_C_T, aes(x = general_votes, 
                                         y = amt, 
                                         color = Candidate)) +
  geom_point() +
  geom_smooth() +
  scale_y_log10() +
  scale_x_sqrt() +
  labs(title = "Contributions vs Vote of Clinton and Trump in 2016 Presidential Election",
       x = "Vote",
       y = "Contributions") +
  scale_color_manual(values = c("blue","red")) 
  

figure1<-ggarrange(VS, 
          ggarrange(votes, Vcontribution, ncol = 2), 
          nrow = 2,
          common.legend = TRUE)

annotate_figure(figure1, fig.lab = "Figure 1")
```


## Individual contributors’ occupations for Clinton and Trump
The graph on the left shows the groups with the most individual contributors, regardless of how much money those individuals donated. A group is defined by occupation and party. The graph on the right shows the groups with the greatest total amount of money contributed.
We can see that Trump received money from significantly more individuals than Clinton, but the margin between them for the total amount of money received was much smaller. This suggests that each person who contributed to Trump may have given a smaller amount of money on average.

The only main contributing occupations for which Clinton has more individual supporters than Trump are anesthesiologist, clerk, and national field director.

Both parties are influenced heavily by money given by presidents, but presidents are very few individuals who do not appear on the count graph. The second biggest monetary contributor is retired people, who also make up a significant number of individuals. This may indicate that an older demographic tends to be the most active in donating to elections.

Those who are unemployed also have a surprisingly large number of contributors and amount of contributions. Those who are unemployed contributed nearly equal amounts of money to both parties, but Trump received money from far more unemployed individuals. This may say something about the reasons for unemployment for individuals for each party; Trump has a much larger number of unemployed supporters who contributed small amounts of money, while Clinton has few unemployed supporters who each gave a larger amount. 




```{r}
individuals2 <- individuals %>% 
  select(occupation, cmte_id, transaction_amt, tran_id)

candidates2 <- candidates %>% 
  select(cand_id, cand_pty_affiliation)

contributions2 <- contributions %>% 
  select(cmte_id, cand_id) %>% 
  inner_join(individuals2, by = "cmte_id") %>% 
  inner_join(candidates2, by = "cand_id") %>% 
  group_by(occupation, cand_pty_affiliation) %>% 
  summarize(total = sum(transaction_amt)) %>% 
  filter(cand_pty_affiliation == "DEM" | cand_pty_affiliation == "REP") %>% 
  arrange(desc(total)) %>% 
  head(30)

occupations<-contributions %>% 
  select(cmte_id, cand_id) %>% 
  inner_join(individuals2, by = "cmte_id") %>% 
  inner_join(candidates2, by = "cand_id") %>% 
  filter(cand_id == "P00003392" | cand_id == "P80001571") %>% 
  group_by(cand_id, occupation) %>% 
  summarise(count = n()) %>% 
  mutate(Candidate = ifelse(cand_id == "P00003392", "Clinton", "Trump")) %>% 
  arrange(desc(count)) %>% 
  head(30)

occupation_by_count<-ggplot(occupations, aes(x = occupation, 
                                             y = count, 
                                             fill = Candidate)) +
  geom_col(position = position_dodge(preserve = "single")) +
  theme(axis.text.x = element_text(size = 4),
        axis.text.y = element_text(size = 4)) +
  scale_fill_manual(values = c("blue","red")) +
  labs(x = "Occupation", y = "Count")+coord_flip()

occupation_by_amt<-ggplot(data = contributions2, 
                          mapping = aes(x = occupation, 
                                        y = total, 
                                        fill = cand_pty_affiliation)) +
  geom_col(position = position_dodge(preserve = "single")) +
  theme(axis.text.x = element_text(size = 4),
        axis.text.y = element_text(size = 4))+
  scale_fill_manual(values = c("blue","red"))+
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Occupation", y = "Contribution (dollars)")+coord_flip()

figure2 <- ggarrange(occupation_by_count, occupation_by_amt,
                    ncol = 2, 
                    common.legend=TRUE)
annotate_figure(figure2, 
               top = text_grob("Individual Contributors' Occupations of Clinton and Trump 
                               in the 2016 Presidential Election", 
                               face = "bold"),
               fig.lab = "Figure 2")
```



